---
title: "Google page rock!"
author: "Jennifer Schmidt"
date: "August 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages({
  library(leaflet) # interactive mapping dat
  library(dplyr)
  library(tidyr)
  library(ggplot2) # plotting data
  library(DT) # creates very nice tables for the web
})
```


# Header 1  14.4.1 Load salmon escapement data

You can load the data table directly from the KNB Data Repository, if it isn’t already present on your local computer. This technique
Load salmon data from KNB
Try catch trys something and if there is an error then try something else.
```{r}
# data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/knb.92020.1"
# # data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e"
# esc <- tryCatch(
#     read.csv("data/escapement.csv", stringsAsFactors = FALSE),
#     error=function(cond) {
#         message(paste("Escapement file does not seem to exist, so get it from the KNB."))
#         esc <- read.csv(url(data_url, method = "libcurl"), stringsAsFactors = FALSE)
#         return(esc)
#     }
# )
# 
# head(esc)
```


Prepare the dataset for cacluting medians
```{r}
head(esc)
median_esc <- esc %>% 
  separate(sampleDate, c("Year", "Month", "Day"), sep = "-") %>% 
  # separate takes one of your variables and tokens it out.  Three variable year, month, and day.  It is in order.
  group_by(Species, SASAP.Region, Year, Location) %>% 
  summarize(escapement = sum(DailyCount)) %>% 
  group_by(Species) %>% 
  summarize(median_escapement = median(escapement))

head(median_esc)

```
Now, let’s plot our results:
```{r}
library(ggplot2)
ggplot(median_esc, aes(Species, median_escapement)) +
  geom_col() +
  coord_flip()
```

Now let’s convert the escapement data into a table of just the unique locations:
```{r}
locations <- esc %>% 
  distinct(Location, Latitude, Longitude) %>% 
  drop_na()
```

And display it as an interactive table:
```{r}
datatable(locations)
```

<<<<<<< HEAD
Then making a leaflet map is (generally) only a couple of lines of code:
```{r}
leaflet(locations) %>% 
  addTiles() %>% 
  addMarkers(~ Longitude, ~ Latitude, popup = ~ Location)
=======
Collaborator is here
Now following rules

```{r cars}
summary(cars)
>>>>>>> df55ad4175116faae8ab087621d05e6c23c2bdad
```

```{r}
locs <- locations %>% mutate(Longitude = abs(Longitude) * -1)

leaflet(locs) %>% 
  addTiles() %>% 
  addMarkers(~ Longitude, ~ Latitude, popup = ~ Location)
```

First, we use makeIcon to create a local version of the icon symbols to be plotted on the map:

# Use a custom marker so Leaflet doesn't try to grab the marker images from 
# its CDN (this was brought up in 
# https://github.com/NCEAS/sasap-training/issues/22)

```{r}
markerIcon <- makeIcon(
  iconUrl = "https://cdnjs.cloudflare.com/ajax/libs/leaflet/1.3.1/images/marker-icon.png",
  iconWidth = 25, iconHeight = 41,
  iconAnchorX = 12, iconAnchorY = 41,
  shadowUrl = "https://cdnjs.cloudflare.com/ajax/libs/leaflet/1.3.1/images/marker-shadow.png",
  shadowWidth = 41, shadowHeight = 41,
  shadowAnchorX = 13, shadowAnchorY = 41
)
```
and then we use that markerIcon explictly when we call leaflet to draw the map:
```{r}
leaflet(locs) %>% 
  addTiles() %>% 
  addMarkers(~ Longitude, ~ Latitude, popup = ~ Location, icon = markerIcon)
```


